---
title: "NBA Data Cleaning"
author: "Isaac"
format: html
---

## Purpose

This script cleans and organizes raw NBA games, city location, and city ranking
data to make three analysis datasets:

1. Player level game performance and rest stats  
2. Game level geographic information which are city, state, latitude, and longitude  
3. City level rankings from the World’s Best Cities report  

All outputs are written as clean CSV files to the `output/` directory.

## Load packages

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(httr2)
library(jsonlite)
```

## Cleaning Process for Game Stats

```{r}


## raw game logs


tatum_raw  <- readRDS("data/imported_data/tatum_game_log_raw.rds")
murray_raw <- readRDS("data/imported_data/murray_game_log_raw.rds")
grimes_raw <- readRDS("data/imported_data/grimes_game_log_raw.rds")


## column names


tatum_raw  <- janitor::clean_names(tatum_raw)
murray_raw <- janitor::clean_names(murray_raw)
grimes_raw <- janitor::clean_names(grimes_raw)


## player identifiers


tatum_raw  <- tatum_raw  |> dplyr::mutate(player = "Jayson Tatum")
murray_raw <- murray_raw |> dplyr::mutate(player = "Dejounte Murray")
grimes_raw <- grimes_raw |> dplyr::mutate(player = "Quentin Grimes")


## combine players into dataset


games_raw <- dplyr::bind_rows(
  tatum_raw,
  murray_raw,
  grimes_raw
)

##  cleaned analysis dataset


games_clean <- games_raw |>

  # only rows where player had minutes
  dplyr::filter(!is.na(mp), mp != "") |>

  # select  columns
  dplyr::select(
    player,
    date,
    opp,
    mp,
    pts,
    home_away = x
  ) |>

  # clean and create variables
  dplyr::mutate(

    # parse
    date = as.Date(
      lubridate::parse_date_time(
        date,
        orders = c("mdy", "ymd", "b d, Y", "a, b d, Y")
      )
    ),

    # convert points
    pts = as.numeric(pts),

    # convert minutes
    mp = dplyr::if_else(
      stringr::str_detect(mp, ":"),
      as.numeric(lubridate::ms(mp)) / 60,
      as.numeric(mp)
    ),

    # points per minute
    ppm = pts / mp
  )

games_clean

games_final <- games_clean |>

  # remove NA games
  dplyr::filter(!is.na(mp), !is.na(date)) |>
  dplyr::arrange(player, date) |>
  dplyr::group_by(player) |>
  dplyr::mutate(
    days_rest = as.numeric(date - dplyr::lag(date)),
    back_to_back = days_rest == 1
  ) |>
  dplyr::ungroup()

games_final


##  save clean dataset

dir.create(here::here("data", "cleaned_data"), recursive = TRUE)

saveRDS(
  games_final,
  here::here("data", "cleaned_data", "games_final_clean.rds"))
```

## Cleaning Process for City Data from World's Best Cities
```{r}
wbc_raw <- readRDS(
  here("data", "imported_data", "worlds_best_cities_raw.rds")
)

wbc_clean <- wbc_raw |>
  select(
    ranking,
    city,
    state,
    population,
    highlighted_rank_1,
    highlighted_category_1,
    highlighted_category_2
  )

wbc_clean |> slice(1:5)

# save as csv
dir.create(here("data", "cleaned_data"), recursive = TRUE)
write_csv(
  wbc_clean,
  here("data", "cleaned_data", "worlds_best_cities_clean.csv")
)
```

## Create Game Location Dataset
```{r}
games_final <- readRDS(
  here("data", "cleaned_data", "games_final_clean.rds")
)

team_cities <- readRDS(
  here("data", "imported_data", "team_cities_raw.rds")
)

# join city and state
games_with_location <- games_final |>
  left_join(
    team_cities,
    by = c("opp" = "team")
  ) |>
  rename(
    state = region
  )

games_with_location |> slice(1:5)

cities_to_geocode <- games_with_location |>
  select(city, state) |>
  distinct()

cities_to_geocode |> slice(1:5)

# function to geocode city 
geocode_city <- function(city, state) {
  Sys.sleep(1)  
  req <- request("https://nominatim.openstreetmap.org/search") |>
    req_url_query(
      q = paste(city, state),
      format = "json",
      limit = 1
    ) |>
    req_user_agent("DSC210-NBA-Project")

  result <- tryCatch(
    {
      resp <- req |> req_perform()
      data <- resp |> resp_body_string() |> fromJSON()
      if (length(data) == 0) {
        tibble(lat = NA_real_, lon = NA_real_)
      } else {
        tibble(
          lat = as.numeric(data$lat[1]),
          lon = as.numeric(data$lon[1])
        )
      }
    },
    error = function(e) {
      tibble(lat = NA_real_, lon = NA_real_)
    }
  )
  result
}

# run API per city
geo_results <- cities_to_geocode |>
  rowwise() |>
  mutate(geo = list(geocode_city(city, state))) |>
  unnest(geo) |>
  ungroup()

geo_results |> slice(1:5)

## join lat/lon  to games_with_location

games_with_location_final <- games_with_location |>
  left_join(
    geo_results,
    by = c("city", "state")
  )

games_with_location_final |> slice(1:5)

# save csv
write_csv(
  games_with_location_final,
  here("data", "cleaned_data", "games_with_location_clean.csv")
)

```
```{r}
## final output csvs
dir.create(here("output"), recursive = TRUE)

write_csv(
  games_final,
  here("output", "games_final.csv")
)

write_csv(
  games_with_location_final,
  here("output", "games_with_location.csv")
)

write_csv(
  wbc_clean,
  here("output", "worlds_best_cities.csv")
)
```

