---
title: "NBA Data Import"
author: "Isaac"
format: html
---

## Purpose

This collects raw NBA data needed to show how travel and rest impact these three player's performance
during the 2024–2025 season. All data imported here is saved to `data/imported_data/` as uncleaned
files and will be processed later in `cleaning.qmd`.

```{r}
library(here)
library(rvest)
library(httr2)
library(jsonlite)
library(readr)

dir.create(here::here("data", "imported_data"), recursive = TRUE)

```

## Import Tatum  game log

Below is Jayson Tatum's raw game log table scraped from Basketball Reference.
```{r}
url <- "https://www.basketball-reference.com/players/t/tatumja01/gamelog/2025"

tatum_page <- read_html(url)

tatum_tables <- html_elements(tatum_page, "table")

tatum_game_log <- html_table(tatum_tables[[8]])

tatum_game_log

saveRDS(
  tatum_game_log,
  here::here("data", "imported_data", "tatum_game_log_raw.rds")
)
```
## Import  Murray game log

Below is Dejounte Murray's raw game log table scraped from Basketball Reference.

```{r}

url <- "https://www.basketball-reference.com/players/m/murrade01/gamelog/2025"

murray_page <- read_html(url)

murray_tables <- html_elements(murray_page, "table")

murray_game_log <- html_table(murray_tables[[8]])

murray_game_log

saveRDS(
  murray_game_log,
  here::here("data", "imported_data", "murray_game_log_raw.rds")
)
```
## Import Grimes game log

Below is Quentin Grimes's raw game log table scraped from Basketball Reference.

```{r}

url <- "https://www.basketball-reference.com/players/g/grimequ01/gamelog/2025"

grimes_page <- read_html(url)

grimes_tables <- html_elements(grimes_page, "table")

grimes_game_log <- html_table(grimes_tables[[8]])

grimes_game_log

saveRDS(
  grimes_game_log,
  here::here("data", "imported_data", "grimes_game_log_raw.rds")
)
```
## API import geocode

I made an API request to openstreetmap to geocode NBA city coordinates
```{r}
# API request
geo_req <- request("https://nominatim.openstreetmap.org/search") |>
  req_url_query(
    q = "Boston, MA",
    format = "json",
    limit = 1
  ) |>
  req_user_agent("DSC210-NBA-Project")

# request
geo_resp <- geo_req |> req_perform()

# parse 
geo_data <- geo_resp |>
  resp_body_string() |>
  fromJSON()

geo_data
saveRDS(
  geo_data,
  here::here("data", "imported_data", "boston_geocode_raw.rds")
)
```
## CSV import

I created this CSV file of NBA cities and imported it here.
```{r}
team_cities <- readr::read_csv(
  "data/imported_data/team_cities.csv",
  show_col_types = FALSE
)

saveRDS(
  team_cities,
  here::here("data", "imported_data", "team_cities_raw.rds")
)
```

